Trends of SDR in Thessaloniki

### 2023-07-02 Article.tex
Commit 2023-07-02 13:17

Cloud ``shrinking'' and ``optical thinning'' in the ``dimming'' period
and a subsequent recovery in the ``brightening'' period over China

aerosols not only strengthen but also weaken the growth of clouds with
respect to their coverage and optical thickness, depending on the levels
of pollution and the associated amounts of aerosols as postulated in a
conceptual framework (Wild 2009a, 2012, Yang et al 2012).

@@

The shortwave downward solar irradiance (SDR) at the Earth's surface
plays a significant role, on Earths climate. Changes of the SDR can be
related to major changes on factors affecting the radiative forcing
balance and human activities. Multiple studies attempt to evaluate the
phenomenon and the potential causes
\citep{Ohvril2009, Schwarz2020, Wild2009, Wild2012, Xia2007, Zerefos2009}.
Although, there is regional variation of SDR, with both positive and
negative trends, most of the studies agree that the major factors are
the interaction of aerosols and clouds.

In this study, we examine the trends of SDR with ground-based
measurements at Thessaloniki for the period 1993 - 2023, as derived from
a CM-21 pyranometer. We reevaluate and extend the dataset used by
\citet{Bais2013}, applying a different algorithm for the identification
of clear-/cloud-sky instances, and we derive the radiation trends for
this period under different sky conditions (all-sky, clear-sky and
cloud-sky).

Site location description and aerosols effect

@@

The SDR is equivalent to the whole sky Global radiation, also refered as
Global Horizontal Irradiance (GHI), and was obtained with a horizontal
leveled, CM-21 pyranometer. SDR data spans the period of 1993-04-12 to
2023-05-31. In some algorithms we used as auxiliary data, the direct
beam radiation (DNI) for comparisons, obtained by a CHP-1 pyrheliometer,
with data availability from 2016-04-01 to 2023-05-31.

@@

Due to a significant measurement uncertainty near, the horizon, we have
to exclude all measurements with SZA greater than \(85^\circ\).
Moreover, due to some obstructions around the site (hills and
buildings), we excluded data with Azimuth angle between \(35^\circ\) and
\(120^\circ\) with SZA greater than \(80^\circ\). On the latter
instances, Sun is systematically, not visible by the instrument's
location. To make the measurements comparable throughout the dataset, we
adjusted all 1-minute radiometric values to the mean Sun - Earth
distance. Subsequently, we made all measurements relative to the Total
Solar Irradiance (TSI) at \(1 \text{au}\), in order to compensate for
the Sun's intensity variability, using a homogenized time series of
satellite TSI observations. The TSI data we use, is a combination of
data from NOAA \citep{Coddington2005} (for 1993-04-12 - 2023-03-31) and
adjusted data from \citet{TSIStsi} (for 2023-03-31 - 2023-05-31). As a
result, we can present all radiation data as a fraction of Sun's TSI.

@@

\caption{Anomaly (\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for all-sky conditions.}\label{fig:trendALL}
\end{figure}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2048}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1928}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2169}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3855}}@{}}
\caption{\label{tab:trendtable}Trends of daily means by sky conditions.
}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend statistical signif. {[}\%{]}
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend statistical signif. {[}\%{]}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
0.381 & 3.479e-21 & All sky cond. & 100 \\
0.3605 & 1.068e-16 & Clear sky cond. & 100 \\
0.3472 & 8.968e-09 & Cloudy cond. & 100 \\
\end{longtable}

@@

{\centering \subfloat[Trend distribution for all-sky conditions.\label{fig:szatrends-1}]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-13} }\subfloat[Trend distribution for cloud-sky conditions.\label{fig:szatrends-2}]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-21} }

@@

\caption{Distribution of the SDR's long term trends by SZA.}\label{fig:szatrends}
\end{figure}

@@

\caption{Trends by season for all condition. Displaying monthly means of daily means.}\label{fig:seasonalALL}
\end{figure}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2361}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2222}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1250}}@{}}
\caption{\label{tab:trendseasontable}Trends of daily means by sky
conditions for the seasons of the year. (continued
below)}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Season
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Season
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
0.7215 & 0.001209 & All sky cond. & Winter \\
0.1679 & 0.1083 & All sky cond. & Spring \\
0.128 & 0.08705 & All sky cond. & Summer \\
0.4239 & 0.006659 & All sky cond. & Autumn \\
0.8715 & 0.000269 & Clear sky cond. & Winter \\
0.09719 & 0.3923 & Clear sky cond. & Spring \\
-0.02369 & 0.7972 & Clear sky cond. & Summer \\
0.3583 & 0.02996 & Clear sky cond. & Autumn \\
-0.02581 & 0.9335 & Cloudy cond. & Winter \\
0.3557 & 0.06914 & Cloudy cond. & Spring \\
0.442 & 0.004988 & Cloudy cond. & Summer \\
0.4776 & 0.01955 & Cloudy cond. & Autumn \\
\end{longtable}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 0\tabcolsep) * \real{0.4444}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend statistical signif. {[}\%{]}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
99.88 \\
89.17 \\
91.29 \\
99.33 \\
99.97 \\
60.77 \\
20.28 \\
97 \\
6.646 \\
93.09 \\
99.5 \\
98.04 \\
\end{longtable}

@@

{\centering \subfloat[one plot\label{fig:cumsum-1}]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-1} }\subfloat[the other one\label{fig:cumsum-2}]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-9} }

@@

\caption{two plots}\label{fig:cumsum}
\end{figure}

@@

Our result for all-sky condition (\(0.38\%/year\)), reaffirm the
previous results of \citet{Bais2013} for the site. The increase of this
trend, shows that the phenomenon and probably the causes have been
amplified.

\ldots.. Also, the trend of \(0.35\%/year\) for cloud-sky condition
indicate the major part that clouds play on the \ldots..

>>>>

@@

The shortwave downward solar irradiance (SDR) at Earth's surface play a
significant role, on it's climate. Changes of the SDR can be related to
changes on Earth's energy budget, the mechanisms of climate change, and
water and carbon cycle \citep{Wild2009}. Can also affect, solar and
agricultural production, and living organisms. Studies of SDR
variability, have identified some distinct SDR trends on different
regions of the world and for different time periods. The term
`brightening' is generally used to describe periods of positive SDR
trend, and `dimming' for negative trend. There are many cases on the
long term records of irradiance, showing a systematic change of SDR's
trend slope, occurring roughly at the last decades of the 20th century.
On multiple station in China, a dimming period was reported until about
2000, followed by a brightening period \citep{Yang2021}. A similar
pattern was identified, with the breaking point around 1980, for
stations in Central Europe \citep{Wild2021} and Brazil
\citep{Yamasoe2021}. Also, on global scale, an AI aided continental
level spatial analysis, with data from multiple station, reach similar
conclusions for the above regions and for the global trend
\citep{Yuan2021}.

There is a consensus, among researchers, that major factors of SDR
attenuation is the interaction with atmospheric aerosols and clouds.
Those interactions, among other factors, have been analysed with models
\citep{Li2016, Samset2018}, showing the existence of feedback mechanisms
between the two. Similar finds, have been showed in observational data
\citep[ and references
therein]{Schwarz2020, Ohvril2009, Zerefos2009, Xia2007}.

Due to the variability of the phenomenon, and its contributing factors,
there is a need to further investigate and monitor SDR, in different
sites, to estimate its magnitude, and its relation to the local
conditions. In this study, we examine the trends of SDR, with
ground-based measurements at Thessaloniki, Greece for the period 1993 to
2023, as derived from a CM-21 pyranometer. We reevaluated and extended
the dataset used by \citet{Bais2013}, applying a different algorithm for
the identification of clear-/cloud-sky instances
\citep{Reno2016, Reno2012a}, and we derive the radiation trends for the
period 1993 to 2023, under different sky conditions (all-sky, clear-sky
and cloud-sky).

@@

Location

The SDR data used in this study cover the period from January 1993 to
March 2011, and were measured with a Kipp \& Zonen CM-21 pyranometer
operating contin- uously at the Laboratory of Atmospheric Physics of the
Aristotle University of Thessaloniki (\$40\^{}\circ\$38''N,
\$22\^{}\circ\$57''E, ∼80 m above sea level). The monitoring site is
located near the city centre. The horizon is clear of obstruc- ◦ ◦ tions
except for the azimuth angles between 35 and 120where obstacles of up to
∼10◦ elevation block the direct irradiance in the morning for SZA larger
than ∼80◦ . These data points were excluded from the analysis, as

Instruments and callibrations

relation to aerosols

Beijing (40 N, 116 E), which are representative of regions with high
aerosol load. Northern Greece has been selected because it is well known
to be located at the crossroads of aerosols and various pollutants
pathways originating in central and Eastern Europe (Lelieveld et al.,
2002; Zerefos et al., 2002). Thessa- loniki, the major city in northern
Greece, lies along the exit of air pollutants from Europe, in the path
of the prevailing long-range transport (LRT) trajectories, as we move
from central and east- ern Europe to the Mediterranean (Zerefos et. al.,
2002). On the

The SDR is equivalent to the whole sky Global radiation, also refered as
Global Horizontal Irradiance (GHI), and was obtained with a horizontal
leveled, CM-21 pyranometer. SDR data spans the period of 1993-04-12 to
2023-03-31. In some algorithms we used as auxiliary data, the direct
beam radiation (DNI) for comparisons, obtained by a CHP-1 pyrheliometer,
with data availability from 2016-04-01 to 2023-03-31.

@@

Due to a significant measurement uncertainty near, the horizon, we have
to exclude all measurements with SZA greater than \(85^\circ\).
Moreover, due to some obstructions around the site (hills and
buildings), we excluded data with Azimuth angle between \(35^\circ\) and
\(120^\circ\) with SZA greater than \(80^\circ\). On the latter
instances, Sun is systematically, not visible by the instrument's
location. To make the measurements comparable throughout the dataset, we
adjusted all 1-minute radiometric values to the mean Sun - Earth
distance. Subsequently, we made all measurements relative to the Total
Solar Irradiance (TSI) at \(1 \text{au}\), in order to compensate for
the Sun's intensity variability, using a homogenized time series of
satellite TSI observations. The TSI data we use, comes from NOAA
\citep{Coddington2005}, it overlaps with our data, and is interpolated
to match with the time step of the data. As a result, we can present all
radiation data as a fraction of Sun's TSI.

@@

\caption{Anomaly (\%) of the daily SDR relative to climatological values for 1993 - 2023. The black line shows the long term trend for all-sky conditions.}(\#fig:trendALL)
\end{figure}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2048}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1928}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2169}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.3855}}@{}}
\caption{\label{tab:trendtable}Trends of daily means by sky conditions.
}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend statistical signif. {[}\%{]}
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend statistical signif. {[}\%{]}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
0.3996 & 8.665e-23 & All sky cond. & 100 \\
0.3773 & 7.075e-18 & Clear sky cond. & 100 \\
0.3594 & 3.318e-09 & Cloudy cond. & 100 \\
\end{longtable}

@@

{\centering \subfloat[Trend distribution for all-sky conditions.(\#fig:szatrends-1)]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-13} }\subfloat[Trend distribution for cloud-sky conditions.(\#fig:szatrends-2)]{\includegraphics[width=.35\linewidth]{./images/SzaTrends-21} }

@@

\caption{Distribution of the SDR's long term trends by SZA.}(\#fig:szatrends)
\end{figure}

@@

\caption{Trends by season for all condition. Displaying monthly means of daily means.}(\#fig:seasonalALL)
\end{figure}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2361}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2222}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.2500}}
  >{\centering\arraybackslash}p{(\columnwidth - 6\tabcolsep) * \real{0.1250}}@{}}
\caption{\label{tab:trendseasontable}Trends of daily means by sky
conditions for the seasons of the year. (continued
below)}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Season
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend {[}\%/year{]}
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Trend p-value
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sky condition
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Season
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
0.7215 & 0.001209 & All sky cond. & Winter \\
0.1537 & 0.1498 & All sky cond. & Spring \\
0.128 & 0.08705 & All sky cond. & Summer \\
0.4239 & 0.006659 & All sky cond. & Autumn \\
0.8715 & 0.000269 & Clear sky cond. & Winter \\
0.05881 & 0.6247 & Clear sky cond. & Spring \\
-0.02369 & 0.7972 & Clear sky cond. & Summer \\
0.3583 & 0.02996 & Clear sky cond. & Autumn \\
-0.02581 & 0.9335 & Cloudy cond. & Winter \\
0.4333 & 0.02937 & Cloudy cond. & Spring \\
0.442 & 0.004988 & Cloudy cond. & Summer \\
0.4776 & 0.01955 & Cloudy cond. & Autumn \\
\end{longtable}

\begin{longtable}[]{@{}
  >{\centering\arraybackslash}p{(\columnwidth - 0\tabcolsep) * \real{0.4444}}@{}}
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\centering
Trend statistical signif. {[}\%{]}
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
99.88 \\
85.02 \\
91.29 \\
99.33 \\
99.97 \\
37.53 \\
20.28 \\
97 \\
6.646 \\
97.06 \\
99.5 \\
98.04 \\
\end{longtable}

@@

{\centering \subfloat[one plot(\#fig:cumsum-1)]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-1} }\subfloat[the other one(\#fig:cumsum-2)]{\includegraphics[width=.35\linewidth]{./images/CumulativeDailyCumSum-9} }

@@

\caption{two plots}(\#fig:cumsum)
\end{figure}

@@

Our result for all-sky condition (\(0.4\%/year\)), reaffirm the previous
results of \citet{Bais2013} for the site. The increase of this trend,
shows that the phenomenon and probably the causes have been amplified.

\ldots.. Also, the trend of \(0.36\%/year\) for cloud-sky condition
indicate the major part that clouds play on the \ldots..
